---
title: "GG Plot 111 Report"
author: "Gerald Beta"
date: "01/11/2021"
output:
  html_document:
    keep_md: True
    code_folding: "hide"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


## Report overview


This report demonstrates how ggplot can be used to generate different plots using different approaches. GGplot is a widely known  and used package in the R community. It makes life easy and giving out awesome plots.


We used the following packages to complete the work:

library(tidyverse)

library(janitor)

library(patchwork)

library(viridis)

library(ggrepel)

library(ggridges)

library(here)


## Installing packages

Packages are installed using the "install.package" function and loaded using the "library".


```{r loading packages, include=FALSE}

library(tidyverse)
library(janitor)

library(patchwork)
library(viridis)
library(ggrepel)
library(ggridges)

```

```{r include=FALSE}

library(here)

```


## Adding customized theme

We add our customized theme saved as a function by using the 

```{r}
source(here("functions", "theme_gerry.R"))        ## adding my theme
```



## Adding/ reading data into R


```{r}

chlorophyll_original <- read_csv("data/chlorophyll_a.csv")


phosphorus_original <- read_csv("data/phosphorus_profiles.csv")     #adding data into r


```


## Data cleaning 


We clean data using the clean_names function. It removes errors, capital letter, space/gaps and makes data tidy


```{r data cleaning}

chlorophyll <- chlorophyll_original %>% clean_names()
chlorophyll


phosphorus <- phosphorus_original %>% clean_names()        ## cleaning data
phosphorus

```


## Selecting data colums and Changing data to factor



We can select data column we are interested in using the dplyr :: select function or by simply filtering the data using filter function. Here we used dply select for the assignment. For future plotting and combining data sets from different data, it is important to convert your data colum into factors using the mutate(as.factor) function. We then group the data according to how we need to plot or analyse our data. We used geom jitter to plot our points. It allows us to visualize or edit our color balance. Not forgetting the scale_colour_viridis which smoother our plot colors that it to carter for color blind personals. We used geom_violin which is another form of boxplot but it shows a glimpse of how our data distribution. Lab function was used to remove axis labels and adding expression (suitable for statistics)


Chlorophyll data


```{r}

chlorophyll_data <- chlorophyll %>% mutate(lake_name = as.factor(lake_name)) %>%      ## changing data into factors 
                          dplyr::select(lake_name, chlorophyll_concentration_ug_l, sampling_date) %>%   ## selecting column of interest
                          group_by(lake_name, chlorophyll_concentration_ug_l) %>%
                          
                          ggplot() +
                          geom_jitter(aes(x = lake_name, y = chlorophyll_concentration_ug_l, colour = lake_name), width = 0.1, alpha = 0.8) +
                          geom_violin(aes(x = lake_name, y = chlorophyll_concentration_ug_l, colour = lake_name), width = 0.9, alpha = 0.4, size = 0.5) + #3 other version of box plot   
                          coord_cartesian(ylim = c(8,1000)) +
                          labs(x = "", 
                          y = expression(chlorophyll~concentration~(mu*g~.~L^{-1})),   #adding formulas mathematics 
                              size = 3) + 
                          
                          scale_fill_viridis(discrete = TRUE) +                 ## for color blind people
                          theme_gerry() + theme(legend.position = "none")   ## removing legend on our plots
                          
                          
chlorophyll_data
```



## Selecting data colums and Changing data to factor



We can select data column we are interested in using the dplyr :: select function or by simply filtering the data using filter function. Here we used dply select for the assignment. For future plotting and combining data sets from different data, it is important to convert your data colum into factors using the mutate(as.factor) function. We then group the data according to how we need to plot or analyse our data. We used geom jitter to plot our points. It allows us to visualize or edit our color balance. Not forgetting the scale_colour_viridis which smoother our plot colors that it to carter for color blind personals. We used geom_violin which is another form of boxplot but it shows a glimpse of how our data distribution. Lab function was used to remove axis labels and adding expression (suitable for statistics)

phosphorus data 1 using filter


```{r}

phosphorus_plot1 <- phosphorus %>% filter(sample_depth_m > 0) %>%
                arrange(desc(sample_depth_m)) %>%
                mutate(lake_name = as.factor(lake_name)) %>%
  
                ggplot() +
                geom_violin(aes(x = lake_name, y = sample_depth_m, colour = lake_name), alpha = 0.5, width = 0.2) +
                geom_jitter(aes(x = lake_name, y = sample_depth_m, colour = lake_name), width = 0.2) +
                     
                scale_fill_viridis(discrete = TRUE) +
                theme_gerry() + 
                theme(legend.position = "none") +
                labs(x = "",
                y = expression(total~phosphorus~(mu*g~.~L^{-1})))

phosphorus_plot1

```



phosphorus data 2 using mutate



```{r}

phosphorus_data_2 <- phosphorus %>% mutate(lake_name = as.factor(lake_name)) %>%
            dplyr::select(lake_name, sampling_date, total_phosphorus_mg_l) %>%
            group_by(lake_name, sampling_date) %>%
            summarise(total_phosphorus_mg_l = min(total_phosphorus_mg_l, na.rm = TRUE)) %>%
            ggplot() +
            geom_jitter(aes(x = lake_name, y = total_phosphorus_mg_l, colour = lake_name), alpha = 0.9, width = 0.1) +
            geom_violin(aes(x = lake_name, y = total_phosphorus_mg_l, colour = lake_name),
                            alpha = 0.4, size = 0.5, width = 0.9) +
            coord_cartesian(ylim = c(0, 0.35), expand = FALSE) +
            labs(x = "",
            y = expression(total~phosphorus~(mu*g~.~L^{-1}))) +
            scale_fill_viridis(discrete = TRUE) +
            theme_gerry() +
            theme(legend.position = "none")
                               
phosphorus_data_2

```

## Combining plots using patchwork package.


Patchwork allows you to combine or merge two or more plots in one plot. It is very useful especially if you are working with multiply data sets. It works almost similar to facet_wrap function but patchwork combines plots using different data sets, with facet wrap its only one data set showing different plots. Patchwork use the symbol + to combine data.

option 1

```{r combining plot 1}

final_plot_1 <- chlorophyll_data + phosphorus_plot1 + plot_layout(nrow = 2, widths = c(3.2)) +     ## using patchwork to + to combine data
                                        plot_annotation(tag_level = c("A", "B", size = 4)) + ## tagging our plots
                                        theme_gerry() +
                                        theme(legend.position = "none") 
                                        
                                  
final_plot_1

```


option 2 


```{r}
final_plot_2 <- chlorophyll_data + phosphorus_data_2 + plot_layout(nrow = 2, widths = c(3.2)) +     ## using patchwork to + to combine data
                                        plot_annotation(tag_level = c("A", "B", size = 4)) + ## tagging our plots
                                        
                                        theme_gerry() +
                                        theme(legend.position = "none")
                                      
                                              
final_plot_2                                       
```




Saving plots as pdf


```{r}

ggsave(here("figures", "final_plot_1.pdf"), final_plot_1, width = 190, height = 120, units = "mm")

ggsave(here("figures", "final_plot_2.pdf"), final_plot_1, width = 190, height = 120, units = "mm")

```



Happy coding !!
